### 2020-07-09 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
final-ish draft of preliminary model runs

### Sample Size for each mortality class
```{r}
#mcf -------------------------------------------------------------------
mortality_count <- mcf %>% 
  group_by(mortality_qual) %>% 
  count()

@@

rdnbr_mcf <- mcf %>% 
  group_by(rdnbr_qualitative) %>% 
  count()

dnbr_mcf <- mcf %>% 
  group_by(dnbr_qualitative) %>% 
  count()

#coulter ------------------------------------------------------------------- 
rdnbr_pico <- pico %>% 
  group_by(rdnbr_qualitative) %>% 
  count()

dnbr_pico <- pico %>% 
  group_by(dnbr_qualitative) %>% 
  count()

@@

# all
ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
  geom_jitter()

#by prefire cover
ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
  geom_jitter() +
  facet_wrap(~prefire)

## use lm to get a rdnbr threshold for the type converter stands? ---------------------

@@

# potential type conversion threshold --------------------------------------------

tc_threshold <- mcf %>% 
  mutate(rdnbr_tc_threshold = case_when(
    rdnbr_2007 <1160 ~ "not_tc",
    rdnbr_2007 >1160 ~ "tc"
  ))

@@

rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                    data = mcf, 
                    sampsize = c('None' = 158, 'Low' = 158, 'Moderate' = 158, 'High' = 158), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

@@

### RF manual: type conversion
Just using for loop to generate several graphs so its easier to compare them and make sure importance is stable

rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                    data = mcf, 
                    sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

```
variable importance is KINDA/MOSTLY stable. There arent big changes between runs, some variables flip 1 space, but some move 3 spaces. Variable importance was more stable when ndvi was included in the model.

@@

>>>>

### Sample Size for each mortality class
```{r}
#mcf -------------------------------------------------------------------
mortality_count <- mcf %>% 
  group_by(mortality_qual) %>% 
  count() %>% 
  rename(manual_mortality=mortality_qual)

@@

rdnbr_mcf <- mcf %>% 
  group_by(rdnbr_qualitative) %>% 
  count() %>% 
  rename(mcf_rdnbr = rdnbr_qualitative)

dnbr_mcf <- mcf %>% 
  group_by(dnbr_qualitative) %>% 
  count() %>% 
  rename(mcf_dnbr = dnbr_qualitative)

#coulter ------------------------------------------------------------------- 
rdnbr_pico <- pico %>% 
  group_by(rdnbr_qualitative) %>% 
  count() %>% 
  rename(pico_rdnbr =rdnbr_qualitative)

dnbr_pico <- pico %>% 
  group_by(dnbr_qualitative) %>% 
  count() %>% 
  rename(pico_dnbr = dnbr_qualitative)

@@

# mcf
ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
  geom_jitter()

##by prefire cover
ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
  geom_jitter() +
  facet_wrap(~prefire)

## use lm to get a rdnbr threshold for the type converter stands? pretty sure thats not what you meant chris ---------------------

@@

# potential type conversion threshold --------------------------------------------
## pretty sure i did NOT followed the correct steps to pick this threshold. 
tc_threshold <- mcf %>% 
  mutate(rdnbr_tc_threshold = case_when(
    rdnbr_2007 <1160 ~ "not_tc",
    rdnbr_2007 >1160 ~ "tc"
  )) %>% 
  mutate(rdnbr_tc_threshold = as.factor(rdnbr_tc_threshold))

tc_threshold_count <- tc_threshold %>% 
  group_by(rdnbr_tc_threshold) %>% 
  count() # not tc = 702, tc = 33

@@

rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi, 
                    data = mcf, 
                    sampsize = c('None' = 158, 'Low' = 158, 'Moderate' = 158, 'High' = 158), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

@@

@@

# Type conversion: manual mortality (classification)

rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi, 
                    data = mcf, 
                    #sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

# Type conversion: rdnbr threshold (classification)

preliminary threshold: 1160
```{r}

rfm <- randomForest(rdnbr_tc_threshold ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi, 
                    data = tc_threshold, 
                    #sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
```
ROC and AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(tc_threshold$rdnbr_tc_threshold, rfm$votes[,1])
plot(rfm_roc) # plot comes out a little weird
auc(rfm_roc)
```

PDP
```{r}
#PDP
a <- partialPlot(rfm, as.data.frame(tc_threshold), rain, "not_tc") 
b <- partialPlot(rfm, as.data.frame(tc_threshold), rain, "tc", col="red") 
lines(a, col="green")

a <- partialPlot(rfm, as.data.frame(tc_threshold), slope, "not_tc") 
b <- partialPlot(rfm, as.data.frame(tc_threshold), slope, "tc", col="red") 
lines(a, col="green")

a <- partialPlot(rfm, as.data.frame(tc_threshold), vpd_aug, "not_tc") 
b <- partialPlot(rfm, as.data.frame(tc_threshold), vpd_aug, "tc", col="red") 
lines(a, col="green")

```
ALE
```{r}
#sub data
tc_threshold_ale <- tc_threshold %>% 
  select(rdnbr_tc_threshold, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed, ndvi) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.

train_x <- tc_threshold_ale[-which(names(tc_threshold_ale) == "rdnbr_tc_threshold")]

# predictors 
nottc_predictor <- Predictor$new(rfm, data = tc_threshold_ale, y = tc_threshold_ale$rdnbr_tc_threshold, type = "prob", class="not_tc")
tc_predictor <- Predictor$new(rfm, data = tc_threshold_ale, y = tc_threshold_ale$rdnbr_tc_threshold, type = "prob", class="tc")

#sr
other <- plot(FeatureEffect$new(nottc_predictor, feature = "rain")) + ggtitle("Not Type Converted")
high <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted")
ggarrange(other, high)
